---
title: "Finding GEO Expression Data"
Date: "February 8, 2022"
output: "None"
author: "Abdula Maher"
---

Install the Bioconductor and GEOmetadb packages if they are not installed
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

if (!requireNamespace("GEOmetadb", quietly = TRUE))
  install.packages("GEOmetadb")

library("BiocManager")
library("GEOmetadb")
```

Obtain GEO data and save it as a SQL lite file in the working directory
```{r}
if(!file.exists('GEOmetadb.sqlite'))
  getSQLiteFile()
```

Connect to the newly downloaded GEOmeta data database and examine the available tables
```{r}
con <- dbConnect(SQLite(), 'GEOmetadb.sqlite')

geo_tables <- dbListTables(con)
geo_tables
```
Run SQL queries and examine the data on GEO:
(1) Look at the different technologies on GEO and highlight RNA-seq (classified under high-throughout sequencing)
```{r}
uniq_tech <- dbGetQuery(con, 'select distinct technology from gpl')
mod_table <- cbind(uniq_tech[1:(nrow(uniq_tech)/2), 1],
                   uniq_tech[((nrow(uniq_tech)/2)+1):nrow(uniq_tech),1])
knitr::kable(mod_table, format = "html")
```

(2) Look at the names GEOmetadb at the technologies that are there for human studies
```{r}
num_uniq_tech_human <- dbGetQuery(con, ('select technology,count(*) as num_plat from gpl where organism = "Homo sapiens" group by technology order by num_plat desc'))
colnames(num_uniq_tech_human)[2] <- "Num_Platforms"
dim(num_uniq_tech_human)
knitr::kable(num_uniq_tech_human, format = "html")
```

Find dataset of interest:
RNA-seq data in AML in humans that was created in the last 5 years + specify for datasets with supplementary file (not a RAW file)
```{r}
sql <- paste("SELECT DISTINCT gse.title,gse.gse, gpl.title,",
             "  gse.submission_date,",
             "  gse.supplementary_file",
             "FROM",
             "   gse JOIN gse_gpl ON gse_gpl.gse=gse.gse",
             "   JOIN gpl ON gse_gpl.gpl=gpl.gpl",
             "WHERE",
             "   gse.submission_date > '2017-01-01' AND",
             "   gse.title LIKE '%Acute Myeloid Leukemia%' AND",
             "   gpl.technology = 'high-throughput sequencing' AND",
             "  gpl.organism = 'Homo sapiens'",
             "order by gse.submission_date desc",
             sep=" ")

rs <- dbGetQuery(con, sql)

count_files <- rs$supplementary_fule[grep(rs$supplementary_file,
                                          pattern = "count",ignore.case = TRUE)]

knitr::kable(rs[, 1:4], format = "html")
```

Series of interest = GSE128603 - Obtain data for the series of interest
```{r}
sfiles = getGEOSuppFiles('GSE128603')

fnames = rownames(sfiles)

b2 = read.delim(fnames[1],header=TRUE)
head(b2)
```

